%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 6CCS3PRJ Final Year Individual Project Report
% luke.day@kcl.ac.uk
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[11pt]{informatics-report}
\usepackage{color}
\usepackage[square,sort,comma,numbers]{natbib} %References
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{mathabx}
\usepackage{graphicx}
\usepackage{tikz-qtree}
\usepackage{url}
\usepackage{float}
\usepackage{graphicx}
\usepackage{ stmaryrd }
\usepackage[T1]{fontenc}
\usepackage{inconsolata}
\usepackage{subcaption}


\lstset{
  basicstyle=\footnotesize\tt,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  language=Java,                 % the language of the code
  keywordstyle=\bf,
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  tabsize=2                       % sets default tabsize to 2 spaces
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Front Matter - project title, name, supervisor name and date
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{6CCS3PRJ Final Year\\\vspace{0.2cm}Automated Timeline Extraction}
\author{Oliver Philip H\"ohn}
\studentID{1426248}
\supervisor{Dr Jeroen Keppens}

\date{\today}

\abstractFile{FrontMatter/abstract.tex}
\ackFile{FrontMatter/acknowledgements.tex} %Remove line if you do not want acknowledgements

\begin{document}
\createFrontMatter
\onehalfspacing
\tableofcontents
\doublespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Report Content
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% You can write each chapter directly here or in a separate .tex file and use the include command.

\chapter{Introduction}
This is one of the most important components of the report. It should begin with a clear statement of what the project is about so that the nature and scope of the project can be understood by a lay reader. It should summarise everything that you set out to achieve, provide a clear summary of the project's background and relevance to other work, and give pointers to the remaining sections of the report, which will contain the bulk of the technical material.
-What is Project About (incl scope)? What is Aim? Bakground? Relevance to other works? (Pointers to other sections?)
\par The project aims to facilitate the understandment of a substantial number of documents (espeically law related) through a graphical representation. When a law professional is tasked with a law case, they are expected to fully understand the overall structure and occurrence of the events described in the documents. However, when a large collection of documents are involved, this task can be both cumbersome for the employees and  financially expensive for the employer. As the documents tend to be extensive and complex, requiring many working hours to understand the underlying story, thus not allowing the professional to progress to the next stage of their work.
\section{Project Scope}
\par As the system requires analyzing, processing and graphically representing the information in the documents, the main areas of the project are Natural Language Processing (NLP), Data Processing and Data Representation. There relevance to other works will be discussed in the Background Chapter.
\section{Objectives}
//what is an event (summary, size, subjects, date, etc)
\par The Objectives are to produce a system that is effective, efficient and simple to use. The system should require as input a selection of documents written in correct English (i.e. in natural language) and produce a timeline with the events described in the documents.
\par Since the user do not have to be computer scientists or have any technological knowledge, but rather are expected to be law professionals, the system should be approachable for them. Therefore the visibility of what the system is doing, and what the users options are is important. 
\par The timeline should be self-expanatory. The user should understands which events happened during which time periods, and know from the information provided what occurred during an event. 
\par The system should produce responses in an appropriate time based on the input, and allow the user to rectify where the system has made mistakes. 
\par The system should allow users to change the input settings, such as the length of the summary, reference points or how many processors can the program use in parallel (to limit/improve the performance), to allow the user to see how the timeline changes with different input parameters.
\par The user may consider the need to save the timeline for later use, or further analyzation. Thereby, providing tools to save the data of the timeline as a PDF or as a JSON in the user's system would benefit them. Producing a JSON output would allow the system to be used by 3rd-parties, that may provide their own graphical user interface (UI) using the system to only process the documents; or further process the events identified by the system.
\section{Report Structure}
\par Following this chapter is a discussion of the background related to this project. Ths is then followed by a formal presentation of the requirements and specification of the project. Which is then followed by the design architecture and patterns used in the system. Followed by the implementation, testing and evaluation of the system. During the evaluation, it will be discussed to what extend the requirements have been met. In the final chapter, the project will be concluded and improvements for future work will be discussed.


\include{Chapters/Background}
\include{Chapters/Body}
\include{Chapters/RequirementsSpecification}
\include{Chapters/Design}
\include{Chapters/Implementation}
\include{Chapters/ProfessionalIssues}
\include{Chapters/Evaluation}
\include{Chapters/Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% References
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{references}
\addcontentsline{toc}{section}{Bibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendices
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\include{Appendices/appendix}
\include{Appendices/UserGuide}
\include{Appendices/SourceCode}
\end{document}
